---
title: "CARPALL standard analysis"
format: html
---

```{r}
setwd("C:/Users/dan_m/OneDrive - University College London/Documents/CARPALL_scRNAseq_raw")
```

```{r}
library(Seurat)
library(SeuratObject)
library(RColorBrewer)
library(harmony)
library(clustree)
library(tidyr)
library(viridis)
library(dplyr)
library(data.table)
library(scCustomize)
library(SeuratWrappers)
library(gridExtra)
library(cowplot)
library(patchwork)
library(ggsci)
library(ggthemes)
library(ggplot2)
library(DoubletFinder)
```

### get ist of files from the working directory. i will keep certain files (as some are not going to be analysed, as they were aligned/sequenced with different settings)

```{r}
files <- list.files(path = getwd())
files <- files[c(19:115)]
files
data <- list()
```

### there are more samples in this folder than samples listed in the metadata, so i need to find out what the extra samples are and if they should be kept
### metadata comes from Nathan's github
### the metadata in the paper is laid out differently, so difficult to reorganise (and might be different in general)
```{r}
metadata <- readxl::read_xlsx("metadata_for_analysis.xlsx")
```

```{r}
filenames <- sub(".*(CAR-T[0-9]+).*", "\\1", files)
different_files <- setdiff(filenames, metadata$GEX_Sample)
# based on this, there are 9 samples that have (what appears) to be GEX files, but in the metadata these should only be TCR sequenced files
# until i establish what these are, it's probably safest to exclude them
```

```{r}
files_to_keep <- filenames %in% metadata$GEX_Sample
files <- files[files_to_keep]
```

### create seurat objects from chosen samples
```{r}
for (i in files) {
  # Read 10X data
  seurat_data <- Read10X(data.dir = i)
  # Handle both list (multi-assay) and matrix (single-assay) formats
  if (is.list(seurat_data)) {
    # Prefer "Gene Expression" if present
    if ("Gene Expression" %in% names(seurat_data)) {
      counts <- seurat_data[["Gene Expression"]]
    } else {
      warning(paste("No 'Gene Expression' assay found for", i, "- skipping this sample"))
      next  # skip this sample
    }
  } else {
    # Single assay: assume it's gene expression
    counts <- seurat_data
  }
  
  # Rename cells to include sample ID to avoid barcode collision
  colnames(counts) <- paste(i, colnames(counts), sep = "_")
  # Create Seurat object
  data[[i]] <- CreateSeuratObject(counts = counts, 
                                  min.features = 100, 
                                  project = i)
}

sample_names <- names(data)

#give each sample its own name
for (i in seq_along(data)) {
  data[[i]]$orig.ident <- sample_names[i]
}

#merge all samples into one seurat object called sepsis
carpall <- merge(
  x = data[[1]],
  y = data[-1],
  add.cell.ids = sample_names,
  project = "CARPALL"
)
```

### remove temporary data object and add sample column
```{r}
rm(data)
table(carpall$orig.ident)
carpall$sample <- carpall$orig.ident
table(carpall$sample)
carpall
```

### each sample is a layer currenty, so we join them
```{r}
carpall <- JoinLayers(carpall)
carpall
```

### DOUBLET DETECTION
### samples with fewer than 50 cells will cause scdblfinder to fail, so should be removed. also in general so fews cells is unreliable for analysis - this removed 2 samples
```{r}
library(scDblFinder)
library(BiocParallel)
sce <- as.SingleCellExperiment(carpall)

cell_counts <- table(carpall$sample)
bad_samples <- names(cell_counts[cell_counts < 50])
sce <- sce[, !(sce$sample %in% bad_samples)]

dim(sce)

sce <- scDblFinder(sce, samples="sample", dbr.sd=1)
table(sce$scDblFinder.class)
table(sce$scDblFinder.score)
```

### add doublet metrics to seurat object
```{r}
carpall$scDblFinder.class <- NA
carpall$scDblFinder.score <- NA

common_cells <- intersect(colnames(carpall), colnames(sce))

carpall$scDblFinder.class[common_cells] <- sce$scDblFinder.class
carpall$scDblFinder.score[common_cells] <- sce$scDblFinder.score
```

### remove sce object for memory
```{r}
rm(sce)NormalizeData
```

### FILTERING
```{r}
carpall[["percent.mt"]] <- PercentageFeatureSet(carpall, pattern = "^MT-")
VlnPlot(carpall, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0)
```

```{r}
carpall <- subset(carpall, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 20)
```

### FILTERED SAVE/LOAD POINT
```{r}
#saveRDS(carpall, file="CARPALL_filtered.rds", compress=F)
carpall <- readRDS(file="CARPALL_filtered.rds")
```

### PROCESSING
```{r}
carpall <- NormalizeData(carpall)
carpall <- FindVariableFeatures(carpall)
carpall <- ScaleData(carpall, vars.to.regress = c("nCount_RNA", "percent.mt"))
carpall <- RunPCA(carpall)
ElbowPlot(carpall)
```

### PROCESSED SAVE/LOAD POINT
```{r}
#saveRDS(carpall, file="CARPALL_processed.rds", compress=F)
carpall <- readRDS(file="CARPALL_processed.rds")
```

### CLUSTERING
```{r}
carpall <- FindNeighbors(carpall, dims = 1:12, reduction = "pca")
carpall <- RunUMAP(carpall, dims = 1:12, reduction = "pca", seed.use =61)
```

### EMBEDDED SAVE/LOAD POINT
```{r}
#saveRDS(carpall, file="CARPALL_embedded.rds", compress=F)
carpall <- readRDS(file="CARPALL_embedded.rds")
```

### make sample id cleaner and also compatible with metadata
```{r}
carpall$sample_id <- sub(".*(CAR-T[0-9]+).*", "\\1", carpall$sample)
```

### add metadata to seurat object
```{r}
metadata <- as.data.frame(metadata)
rownames(metadata) <- metadata$GEX_Sample
meta_to_add <- metadata[
  carpall$sample_id,
  c("Patient", "Sort", "Source", "Timepoint", "Timepoint_bin")
]

rownames(meta_to_add) <- colnames(carpall)

carpall <- AddMetaData(carpall, meta_to_add)
```

### check numbers look reasonable
```{r}
table(carpall$Patient, carpall$Timepoint_bin)
```

```{r}
DimPlot(carpall, group.by = c("Patient", "Timepoint_bin", "Sort", "Source"))&scale_colour_npg()&theme_void()
```

```{r}
FeaturePlot(carpall, features = c("CAT-scFv", "CD3D", "CD8A", "LYZ", "CD79A", "NCAM1"))&scale_colour_viridis(option = "F", direction = -1)
```