---
title: "Untitled"
format: html
---


```{r}
library(Seurat)
library(SeuratObject)
library(RColorBrewer)
library(harmony)
library(clustree)
library(tidyr)
library(viridis)
library(dplyr)
library(data.table)
library(scCustomize)
library(gridExtra)
library(cowplot)
library(patchwork)
library(ggsci)
library(ggthemes)
library(ggplot2)
library(anndataR)
library(cartography)
library(DESeq2)
library(clusterProfiler)
library(org.Hs.eg.db)
```

```{r}
catcar <- readRDS("CARPALL_CD3posCATpos.rds")
```

### run processing pipeline again
```{r}
catcar <- NormalizeData(catcar)
catcar <- FindVariableFeatures(catcar)
catcar <- ScaleData(catcar, vars.to.regress = c("nCount_RNA", "percent.mt"))
catcar <- RunPCA(catcar)
ElbowPlot(catcar)
```


### batch correction using harmony
### i'm going to use patient as the batch key as a reasonable stop-gap until i know the true batches

```{r}
catcar <- RunHarmony(catcar, group.by.vars = c("Patient"))
```

### run clustering and embedding
```{r}
catcar <- FindNeighbors(catcar, dims = 1:12, reduction = "harmony")
catcar <- RunUMAP(catcar, dims = 1:12, reduction = "harmony", seed.use =61)
resolutions <- c(0.5,0.6,0.7,0.8,0.9,1)
for (i in resolutions) {
    catcar <- FindClusters(catcar, resolution = i, cluster.name = paste0("resolution_",i))
}
```

### SAVE/LOAD POINT
```{r}
#saveRDS(catcar, file = "car_only_clustered.rds", compress = F)
catcar <- readRDS("car_only_clustered.rds")
```

### some QC umaps of stress, apoptosis and cell cycle genes
```{r}
stress_genes <- c( "FOS",
    "HSPA1A", "HSPA1B", "HSPB1")

apoptosis_genes <- c("CASP3", "CASP6", "FAS", "MAP2K1")

stress_plot <- FeaturePlot(catcar, features = stress_genes)&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
ggsave(plot=stress_plot, file="catcar_stress_genes.png", dpi = 300, height = 6, width = 6)

apoptosis_plot <- FeaturePlot(catcar, features = apoptosis_genes)&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
ggsave(plot=apoptosis_plot, file="catcar_apoptosis_genes.png", dpi = 300, height = 6, width = 6)
```


### add remission information
```{r}
catcar$Remission <- NA

catcar$Remission[catcar$Patient %in% c("P01","P02","P03","P05")] <- "Remission"
catcar$Remission[catcar$Patient %in% c("P06","P08","P09","P10","P11","P13")] <- "Relapse"

table(catcar$Remission)
```

### order some groups to make more sense
```{r}
catcar$Timepoint_bin <- factor(catcar$Timepoint_bin, levels = c("Product", "Early", "Mid", "Late"))
```

### exploratory plots
```{r}
p <-DimPlot(catcar, group.by = c("Patient", "Timepoint_bin", "Source", "Remission"))&scale_colour_npg()&
    theme_void() &
    guides(color = guide_legend(override.aes = list(size = 6))) &
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
p
ggsave(plot=p, filename="car_only_umap_metadata.png", dpi=300, height = 10, width = 12, bg="white")
```

```{r}
p <- FeaturePlot(catcar, features = c("CAT-scFv", "CD3D", "CD8A", "CD4", "LYZ", "CD79A", "NKG7", "NCAM1", "GZMB"))&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
p
ggsave(plot=p, filename = "car_only_key_marker_featureplot.png", dpi=300,height=10,width=12)
```

```{r}
DimPlot(catcar, group.by = c("celltypist_high_label"))&theme_void()
FeaturePlot(catcar, features = c("celltypist_low_score", "celltypist_high_score"))&
    theme_void()&
    scale_colour_viridis(option = "F", direction = -1)
```

```{r}
DimPlot(catcar, group.by = c("resolution_0.5"))&
    theme_void()&scale_colour_d3("category20")
```


### clean up detailed celltypist labels to bin rare populations into "other"
### any population less than 1% of cells (1860 cells) is labelled as other
```{r}
celltype_counts <- table(catcar$celltypist_low_label)

# identify labels with <100 cells
small_types <- names(celltype_counts[celltype_counts < 1860])

# create cleaned column
catcar$celltypist_cleaned <- ifelse(
  catcar$celltypist_low_label %in% small_types,
  "Other",
  catcar$celltypist_low_label
)
```

### make unified colour palette for plots
```{r}
celltypes <- levels(factor(catcar$celltypist_cleaned))

cols <- pal_d3("category20")(length(celltypes))

names(cols) <- celltypes
```

```{r}
p <- DimPlot(catcar, group.by = c("celltypist_cleaned"), raster = F)&theme_void()&scale_colour_d3("category20")
p
ggsave(plot=p, filename="car_only_umap_celltypist_cleaned.png", dpi=300, height = 6, width = 8, bg="white")
```

### proportions of celltypist populations
```{r}
p <- Proportion_Plot(catcar, group.by="celltypist_cleaned", split.by="Timepoint_bin")+scale_fill_manual(values = cols)
p
ggsave(plot=p, filename = "catcar_celltypist_cleaned_proportions_by_time.png", dpi=300, height = 6, width = 6)
```

### proportions by outcome
```{r}
p <- Proportion_Plot(catcar, group.by="celltypist_cleaned", split.by="time_outcome")+scale_fill_manual(values = cols)
p
ggsave(plot=p, filename = "catcar_celltypist_cleaned_proportions_by_time_outcome.png", dpi=300, height = 6, width = 8)
```

### PSEUDOBULK DIFFERENTIAL EXPRESSION ANALYSIS
### comparing timepoints of product first
```{r}
Idents(catcar) <- "Timepoint_bin"
bulk_catcar <- AggregateExpression(catcar, assays="RNA", return.seurat=T, group.by = c("Patient", "Timepoint_bin"))
colnames(bulk_catcar)
```

### aggregating messes with metadata columns, so adding back in
```{r}
meta <- data.frame(sample = colnames(bulk_catcar))
meta <- separate(
    meta,
    col = sample,
    into = c("Patient", "Timepoint_bin"),
    sep = "_"
)
rownames(meta) <- colnames(bulk_catcar)
bulk_catcar <- AddMetaData(
    bulk_catcar,
    metadata = meta
)
```

### set up deseq dataset
```{r}
counts <- GetAssayData(
    bulk_catcar,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = bulk_catcar@meta.data,
    design = ~ Patient + Timepoint_bin
)
```


### first running LRT test just to identify genes different at any point over time
```{r}
dds <- DESeq(dds, test = "LRT", reduced = ~ Patient)
res_lrt <- results(dds)
res_lrt_df <- as.data.frame(res_lrt)
sig_lrt <- subset(
    res_lrt_df,
    padj < 0.05
)
```

### im also going to put the overall deg list into a dotplot, then extract the data to get the percentage positive values
### if a gene isn't expressed in at least 25% (subject to chnge) of cells in one timepoint, i'll exclude it as it's likely not that impactful
### ill then the updated list forward
```{r}
all_degs <- rownames(sig_lrt)
pct_exp <- DotPlot(catcar, features = all_degs, group.by = "Timepoint_bin")
pct_exp <- pct_exp$data
pct_exp <- pct_exp[, c(2,3,4)]
pct_exp <- pivot_wider(pct_exp, names_from = "id", values_from = "pct.exp")
#testing 10% to see how many that cuts
degs_filtered <- pct_exp %>%
    filter(if_any(-features.plot, ~ . >= 25))
degs_filtered <- as.data.frame(degs_filtered)
lrt_degs_filtered <- filter(sig_lrt, rownames(sig_lrt) %in% degs_filtered$features.plot)
top25_lrt_degs <- lrt_degs_filtered %>% arrange(padj) %>% slice_head(n=25)
```


### using clustered_dotplot from the scustomise package
```{r}
p <- Clustered_DotPlot(
    catcar,
    features = rownames(top25_lrt_degs),
    group.by = "Timepoint_bin",
    k = 4,
    column_label_size = 10,
    row_label_size=12,
    cluster_ident = F,
    plot_km_elbow = FALSE
)

png(
    filename = "topdegs_clustered_dotplot.png",
    width = 1200,
    height = 2000,
    res = 300
)

draw(p)

dev.off()
```

### the major differences are between the product and all other timepoints, so i will also run deseq2 without the product cells, to specifically look for differences between early, mid and late
```{r}
Idents(bulk_catcar) <- "Timepoint_bin"
no_product <- subset(bulk_catcar, idents = c("Early", "Mid", "Late"))
counts <- GetAssayData(
    no_product,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = no_product@meta.data,
    design = ~ Patient + Timepoint_bin
)
```

```{r}
dds <- DESeq(dds, test = "LRT", reduced = ~ Patient)
res_lrt <- results(dds)
res_lrt_df <- as.data.frame(res_lrt)
sig_lrt <- subset(
    res_lrt_df,
    padj < 0.05
)
top50_lrt_degs <- sig_lrt %>% arrange(padj) %>% slice_head(n=50)
```

### again i will extract genes only present in a certain percentage of cells, as a number of them are very low
```{r}
all_degs <- rownames(sig_lrt)
pct_exp <- DotPlot(catcar, features = all_degs, group.by = "Timepoint_bin")
pct_exp <- pct_exp$data
pct_exp <- pct_exp[, c(2,3,4)]
pct_exp <- pivot_wider(pct_exp, names_from = "id", values_from = "pct.exp")
#testing 10% to see how many that cuts
degs_filtered <- pct_exp %>%
    filter(if_any(-features.plot, ~ . >= 25))
degs_filtered <- as.data.frame(degs_filtered)
lrt_degs_filtered <- filter(sig_lrt, rownames(sig_lrt) %in% degs_filtered$features.plot)
```

```{r}
catcar$Timepoint_bin <- factor(catcar$Timepoint_bin, levels = c("Product", "Early", "Mid", "Late"))
Idents(catcar) <- "Timepoint_bin"
p <- Clustered_DotPlot(
    catcar,
    features = rownames(lrt_degs_filtered),
    group.by = "Timepoint_bin",
    k = 5,
    column_label_size = 10,
    row_label_size=12,
    cluster_ident = F,
    plot_km_elbow = F
)

png(
    filename = "no_product_topdegs_clustered_dotplot.png",
    width = 1200,
    height = 2000,
    res = 300
)

draw(p)

dev.off()
```

### based off of DEGs I'll inspect some specific ones to see if they differ based on remission status

```{r}
catcar$time_outcome <- paste0(catcar$Timepoint_bin, "_", catcar$Remission)
catcar$time_outcome <- factor(catcar$time_outcome, levels = c("Early_Relapse", "Early_Remission", "Mid_Relapse", "Mid_Remission", "Late_Relapse", "Late_Remission", "Product_Remission"))
Idents(catcar) <- "Timepoint_bin"
no_product <- subset(catcar, idents = c("Early", "Mid", "Late"))
```


```{r}
p <- FeaturePlot_scCustom(no_product, features = c("HMOX1"), split.by = "time_outcome", num_columns = 2)&
    scale_colour_viridis(option = "F", direction = -1)&
    theme_void()&NoLegend()&
    theme(plot.title = element_text(size = 12, face = "bold"))
p
ggsave(plot=p, filename = "HMOX1_featureplot_by_remission.png", dpi=300, height = 8, width = 4)
```

```{r}
VlnPlot(no_product, features= "HMOX1", group.by = "time_outcome", pt.size = 0)
```

```{r}
DotPlot(no_product, features = c("IFNAR2", "HMOX1"), group.by = "time_outcome")+scale_colour_viridis(option = "plasma", direction = -1)+coord_flip()
```


### run GSEA on 38 genes

```{r}
genes <- rownames(lrt_degs_filtered)

go_enrich <- enrichGO(
    gene = genes,
    OrgDb = org.Hs.eg.db,
    keyType = "SYMBOL"
)
```

```{r}
go_enrich
```

### no sig enriched terms, so running on original 244 genes (so sig genes but no % expressed threshold)
```{r}
genes <- rownames(sig_lrt)

go_enrich <- enrichGO(
    gene = genes,
    OrgDb = org.Hs.eg.db,
    keyType = "SYMBOL"
)
```

```{r}
enriched_pathways <- go_enrich@result
enriched_pathways <- filter(enriched_pathways, qvalue < 0.05)
```

### now running specific contrasts (for example late vs rest) to get specific signature at a timepoint
### this uses the default wald test
### still excluding product, as the initial LRT told us that this has a strong effect

```{r}
Idents(bulk_catcar) <- "Timepoint_bin"
no_product <- subset(bulk_catcar, idents = c("Early", "Mid", "Late"))
counts <- GetAssayData(
    no_product,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = no_product@meta.data,
    design = ~ Patient + Timepoint_bin
)

dds <- DESeq(dds)

resultsNames(dds)
```

```{r}
res_late_vs_early <- as.data.frame(results(
    dds,
    contrast = c("Timepoint_bin", "Late", "Early"))
)
sig_late_vs_early <- filter(res_late_vs_early, padj < 0.05)
```

### GSEA
```{r}
res_late_vs_early <- res_late_vs_early[!is.na(res_late_vs_early$stat), ]
gene_list <- res_late_vs_early$stat
names(gene_list) <- rownames(res_late_vs_early)
gene_list <- sort(gene_list, decreasing = T)
```

```{r}
library(BiocParallel)
register(SerialParam())
```

```{r}
go_res <- gseGO(
    geneList = gene_list,
    OrgDb = org.Hs.eg.db,
    keyType = "SYMBOL",
    ont = "BP",
    minGSSize = 10,
    maxGSSize = 500,
    pvalueCutoff = 0.05,
    verbose = TRUE
)

go_res <- simplify(
  gsea_res,
  cutoff = 0.7,
  by = "p.adjust",
  select_fun = min,
  measure = "Wang",
  semData = NULL
)
```

```{r}
go_df <- as.data.frame(go_res)
```

```{r}
dotplot(go_res, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

```{r}
# Create dataframe from ranked vector
gene_df <- data.frame(
    SYMBOL = names(gene_list),
    stat = gene_list
)

# Convert symbols to Entrez IDs
gene_conv <- bitr(
    gene_df$SYMBOL,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db
)

# Merge back with statistics
gene_df <- merge(
    gene_df,
    gene_conv,
    by = "SYMBOL"
)

# Remove duplicated Entrez IDs
gene_df <- gene_df[!duplicated(gene_df$ENTREZID), ]

# Create ranked vector for KEGG
kegg_gene_list <- gene_df$stat
names(kegg_gene_list) <- gene_df$ENTREZID

# Sort decreasing
kegg_gene_list <- sort(kegg_gene_list, decreasing = TRUE)
```

```{r}
kegg_res <- gseKEGG(geneList = kegg_gene_list, organism = "hsa")
kegg_df <- as.data.frame(kegg_res)
```

```{r}
dotplot(kegg_res, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

```{r}
top_bottom_df <- kegg_df %>%
    arrange(desc(NES)) %>%
    slice(c(1:10, (n() - 9):n()))
```
```{r}
ggplot(top_bottom_df, aes(x= NES, y = reorder(Description, NES)))+geom_col()+theme_bw()
```