---
title: "Untitled"
format: html
---


```{r}
library(Seurat)
library(SeuratObject)
library(RColorBrewer)
library(harmony)
library(clustree)
library(tidyr)
library(viridis)
library(dplyr)
library(data.table)
library(scCustomize)
library(SeuratWrappers)
library(gridExtra)
library(cowplot)
library(patchwork)
library(ggsci)
library(ggthemes)
library(ggplot2)
library(DoubletFinder)
library(anndataR)
library(cartography)
library(DESeq2)
```

```{r}
catcar <- readRDS("CARPALL_CD3posCATpos.rds")
```

### run processing pipeline again
```{r}
catcar <- NormalizeData(catcar)
catcar <- FindVariableFeatures(catcar)
catcar <- ScaleData(catcar, vars.to.regress = c("nCount_RNA", "percent.mt"))
catcar <- RunPCA(catcar)
ElbowPlot(catcar)
```

### batch correction using harmony
### i'm going to use patient as the batch key as a reasonable stop-gap until i know the true batches

```{r}
catcar <- RunHarmony(catcar, group.by.vars = c("Patient"))
```


### run clustering and embedding
```{r}
catcar <- FindNeighbors(catcar, dims = 1:12, reduction = "harmony")
catcar <- RunUMAP(catcar, dims = 1:12, reduction = "harmony", seed.use =61)
resolutions <- c(0.5,0.6,0.7,0.8,0.9,1)
for (i in resolutions) {
    catcar <- FindClusters(catcar, resolution = i, cluster.name = paste0("resolution_",i))
}
```

```{r}
DimPlot(catcar, group.by = c("Patient", "Timepoint_bin", "Source"), ncol=1)&scale_colour_npg()&theme_void()
```

```{r}
DimPlot(catcar, group.by = c("celltypist_high_label"))&theme_void()
FeaturePlot(catcar, features = c("celltypist_low_score", "celltypist_high_score"))&theme_void()
```

### clean up detailed celltypist labels to bin rare populations into "other"
### any population less than 1% of cells (1860 cells) is labelled as other
```{r}
celltype_counts <- table(catcar$celltypist_low_label)

# identify labels with <100 cells
small_types <- names(celltype_counts[celltype_counts < 1860])

# create cleaned column
catcar$celltypist_cleaned <- ifelse(
  catcar$celltypist_low_label %in% small_types,
  "Other",
  catcar$celltypist_low_label
)
```

```{r}
p <- DimPlot(catcar, group.by = c("celltypist_cleaned"), raster = F)&theme_void()&scale_colour_d3("category20")
p
ggsave(plot=p, filename="car_only_umap_celltypist_cleaned.png", dpi=300, height = 6, width = 8, bg="white")
```

### PSEUDOBULK DIFFERENTIAL EXPRESSION ANALYSIS
### comparing timepoints of product first
```{r}
Idents(catcar) <- "Timepoint_bin"
bulk_catcar <- AggregateExpression(catcar, assays="RNA", return.seurat=T, group.by = c("Patient", "Timepoint_bin"))
colnames(bulk_catcar)
```

### aggregating messes with metadata columns, so adding back in
```{r}
meta <- data.frame(sample = colnames(bulk_catcar))
meta <- separate(
    meta,
    col = sample,
    into = c("Patient", "Timepoint_bin"),
    sep = "_"
)
rownames(meta) <- colnames(bulk_catcar)
bulk_catcar <- AddMetaData(
    bulk_catcar,
    metadata = meta
)
```


### set up deseq dataset
```{r}
counts <- GetAssayData(
    bulk_catcar,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = bulk_catcar@meta.data,
    design = ~ Patient + Timepoint_bin
)
```


### first running LRT test just to identify genes different at any point over time
```{r}
dds <- DESeq(dds, test = "LRT", reduced = ~ Patient)
res_lrt <- results(dds)
res_lrt_df <- as.data.frame(res_lrt)
sig_lrt <- subset(
    res_lrt_df,
    padj < 0.05
)
```

### now running specific contrasts (for example late vs rest) to get specific signature at a timepoint
### this suses the default wald test
### need to also add a new grouping variable to object

```{r}
bulk_catcar$Late_vs_Other <- ifelse(
  bulk_catcar$Timepoint_bin == "Late",
  "Late",
  "Other"
)
bulk_catcar$Late_vs_Other <- factor(bulk_catcar$Late_vs_Other)
```


```{r}
dds <- DESeqDataSetFromMatrix(
  countData = counts,
  colData = bulk_catcar@meta.data,
  design = ~ Patient + Late_vs_Other
)

dds <- DESeq(dds)
```

```{r}
res_late <- results(dds, contrast = c("Late_vs_Other", "Late", "Other"))
res_late_df <- as.data.frame(res_late)
res_late_df$gene <- rownames(res_late_df)
sig_late <- subset(
    res_late_df,
    padj < 0.05
)
```

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)
```


```{r}
gsea_input <- res_late_df$stat
names(gsea_input) <- res_late_df$gene
gsea_input <- na.omit(gsea_input)
gsea_input <- sort(gsea_input, decreasing = T)
```

```{r}
gse <- gseGO(geneList=gsea_input, OrgDb = org.Hs.eg.db, keyType = "SYMBOL")
```