---
title: "Untitled"
format: html
---


```{r}
library(Seurat)
library(SeuratObject)
library(RColorBrewer)
library(harmony)
library(clustree)
library(tidyr)
library(viridis)
library(dplyr)
library(data.table)
library(scCustomize)
library(gridExtra)
library(cowplot)
library(patchwork)
library(ggsci)
library(ggthemes)
library(ggplot2)
library(anndataR)
library(cartography)
library(DESeq2)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
library(enrichplot)
```

```{r}
catcar <- readRDS("CARPALL_CD3posCATpos.rds")
```

### run processing pipeline again
```{r}
catcar <- NormalizeData(catcar)
catcar <- FindVariableFeatures(catcar)
catcar <- ScaleData(catcar, vars.to.regress = c("nCount_RNA", "percent.mt"))
catcar <- RunPCA(catcar)
ElbowPlot(catcar)
```


### batch correction using harmony
### i'm going to use patient as the batch key as a reasonable stop-gap until i know the true batches

```{r}
catcar <- RunHarmony(catcar, group.by.vars = c("Patient"))
```

### run clustering and embedding
```{r}
catcar <- FindNeighbors(catcar, dims = 1:12, reduction = "harmony")
catcar <- RunUMAP(catcar, dims = 1:12, reduction = "harmony", seed.use =61)
resolutions <- c(0.5,0.6,0.7,0.8,0.9,1)
for (i in resolutions) {
    catcar <- FindClusters(catcar, resolution = i, cluster.name = paste0("resolution_",i))
}
```

### SAVE/LOAD POINT
```{r}
#saveRDS(catcar, file = "car_only_clustered.rds", compress = F)
catcar <- readRDS("car_only_clustered.rds")
```

### add remission information
```{r}
catcar$Remission <- NA

catcar$Remission[catcar$Patient %in% c("P01","P02","P03","P05")] <- "Remission"
catcar$Remission[catcar$Patient %in% c("P06","P08","P09","P10","P11","P13")] <- "Relapse"

table(catcar$Remission)
```

### order some groups to make more sense
```{r}
catcar$Timepoint_bin <- factor(catcar$Timepoint_bin, levels = c("Product", "Early", "Mid", "Late"))
```

### exploratory plots
```{r}
p <-DimPlot(catcar, group.by = c("Patient", "Timepoint_bin", "Source", "Remission"))&scale_colour_npg()&
    theme_void() &
    guides(color = guide_legend(override.aes = list(size = 6))) &
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
p
ggsave(plot=p, filename="car_only_umap_metadata.png", dpi=300, height = 10, width = 12, bg="white")
```

```{r}
p <- FeaturePlot(catcar, features = c("CAT-scFv", "CD3D", "CD8A", "CD4", "LYZ", "CD79A", "NKG7", "NCAM1", "GZMB"))&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
p
ggsave(plot=p, filename = "car_only_key_marker_featureplot.png", dpi=300,height=10,width=12)
```

```{r}
DimPlot(catcar, group.by = c("celltypist_high_label"))&theme_void()
FeaturePlot(catcar, features = c("celltypist_low_score", "celltypist_high_score"))&
    theme_void()&
    scale_colour_viridis(option = "F", direction = -1)
```

### want to remove the obvious myeloid cells, so subsetting based on resolution_0.5 and then re-clustering
### whilst here will also filter out predicted doublets

```{r}
DimPlot(catcar, group.by = c("resolution_0.5"))&
    theme_void()&scale_colour_d3("category20")
```

### rename doublet groups and inspect
```{r}
catcar$scDblFinder.class <- factor(catcar$scDblFinder.class, levels = c("1", "2"), labels = c("singlet", "doublet"))
table(catcar$scDblFinder.class)
VlnPlot(catcar, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), group.by = "scDblFinder.class", pt.size = 0)
```

### remove cluster 11 (the myeloid population almost exclusively from patient 6)
```{r}
Idents(catcar) <- "resolution_0.5"
catcar <- subset(catcar, idents = "11", invert = TRUE)
DimPlot(catcar)&
    theme_void()&scale_colour_d3("category20")
```

### remove doublets
```{r}
Idents(catcar) <- "scDblFinder.class"
catcar <- subset(catcar, idents = "singlet")
DimPlot(catcar)&
    theme_void()&scale_colour_d3("category20")
```

### 2nd ROUND OF CLUSTERING
```{r}
catcar <- NormalizeData(catcar)
catcar <- FindVariableFeatures(catcar)
catcar <- ScaleData(catcar, vars.to.regress = c("nCount_RNA", "percent.mt"))
catcar <- RunPCA(catcar)
ElbowPlot(catcar)
catcar <- RunHarmony(catcar, group.by.vars = c("Patient"))
```

```{r}
catcar <- FindNeighbors(catcar, dims = 1:12, reduction = "harmony")
catcar <- RunUMAP(catcar, dims = 1:12, reduction = "harmony", seed.use =61)
resolutions <- c(0.5,0.7, 1)
for (i in resolutions) {
    catcar <- FindClusters(catcar, resolution = i, cluster.name = paste0("resolution_",i))
}
```

### SAVE/LOAD POINT
```{r}
#saveRDS(catcar, file = "car_only_2nd_clustering.rds", compress = F)
catcar <- readRDS("car_only_2nd_clustering.rds")
```

### some QC umaps of stress, apoptosis and cell cycle genes
```{r}
stress_genes <- c( "FOS",
    "HSPA1A", "HSPA1B", "HSPB1")

apoptosis_genes <- c("CASP3", "CASP6", "FAS", "MAP2K1")

stress_plot <- FeaturePlot(catcar, features = stress_genes)&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
ggsave(plot=stress_plot, file="catcar_stress_genes.png", dpi = 300, height = 6, width = 6)

apoptosis_plot <- FeaturePlot(catcar, features = apoptosis_genes)&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
ggsave(plot=apoptosis_plot, file="catcar_apoptosis_genes.png", dpi = 300, height = 6, width = 6)
```

### exploratory plots
```{r}
p <-DimPlot(catcar, group.by = c("Patient", "Timepoint_bin", "Source", "Remission"))&
    scale_colour_d3("category20")&
    theme_void() &
    guides(color = guide_legend(override.aes = list(size = 6))) &
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
p
ggsave(plot=p, filename="car_only_umap_metadata.png", dpi=300, height = 10, width = 15, bg="white")
```

```{r}
p <- FeaturePlot(catcar, features = c("CAT-scFv", "CD3D", "CD8A", "CD4", "LYZ", "CD79A", "NKG7", "NCAM1", "GZMB"))&scale_colour_viridis(option = "F", direction = -1)&theme_void()&NoLegend()&
    theme(
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16),
        plot.title = element_text(size = 18, face = "bold")
    )
p
ggsave(plot=p, filename = "car_only_key_marker_featureplot.png", dpi=300,height=10,width=12)
```

```{r}
DimPlot(catcar, group.by = c("resolution_0.5"))&
    theme_void()&scale_colour_d3("category20")
```

### PROPORTIONS ANALYSIS
```{r}
cluster_props <- catcar@meta.data %>%
  group_by(Patient, Remission, Timepoint_bin, celltypist_cleaned) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Patient, Remission, Timepoint_bin) %>%
  mutate(proportion = (n / sum(n))*100) %>%
  ungroup()
```

```{r}
cluster_props <- filter(cluster_props, Timepoint_bin %in% c("Early", "Mid", "Late"))
```

```{r}
ggplot(cluster_props, aes(x = Timepoint_bin, y = proportion, colour = Remission))+
    facet_wrap(~celltypist_cleaned, scales = "free_y")+
    geom_boxplot(outlier.shape = NA)+
    geom_point(position = position_dodge(width=0.8))+
    theme_bw()+scale_colour_d3("category20")
```


### clean up detailed celltypist labels to bin rare populations into "other"
### any population less than 1% of cells (1860 cells) is labelled as other
```{r}
celltype_counts <- table(catcar$celltypist_low_label)

# identify labels with <100 cells
small_types <- names(celltype_counts[celltype_counts < 1860])

# create cleaned column
catcar$celltypist_cleaned <- ifelse(
  catcar$celltypist_low_label %in% small_types,
  "Other",
  catcar$celltypist_low_label
)
```

### make unified colour palette for plots
```{r}
celltypes <- levels(factor(catcar$celltypist_cleaned))

cols <- pal_d3("category20")(length(celltypes))

names(cols) <- celltypes
```

```{r}
p <- DimPlot(catcar, group.by = c("celltypist_cleaned"), raster = F)&theme_void()&scale_colour_d3("category20")
p
ggsave(plot=p, filename="car_only_umap_celltypist_cleaned.png", dpi=300, height = 6, width = 8, bg="white")
```

### proportions of celltypist populations
```{r}
p <- Proportion_Plot(catcar, group.by="celltypist_cleaned", split.by="Timepoint_bin")+scale_fill_manual(values = cols)
p
ggsave(plot=p, filename = "catcar_celltypist_cleaned_proportions_by_time.png", dpi=300, height = 6, width = 6)
```

### proportions by outcome
```{r}
p <- Proportion_Plot(catcar, group.by="celltypist_cleaned", split.by="time_outcome")+scale_fill_manual(values = cols)
p
ggsave(plot=p, filename = "catcar_celltypist_cleaned_proportions_by_time_outcome.png", dpi=300, height = 6, width = 8)
```

### proportions by patient
```{r}
p <- Proportion_Plot(catcar, group.by="celltypist_cleaned", split.by="Patient")+scale_fill_manual(values = cols)
p
ggsave(plot=p, filename = "catcar_celltypist_cleaned_proportions_by_patient.png", dpi=300, height = 6, width = 8)
```

### CLUSTERING AND ANNOTATION
```{r}
p <- DimPlot(catcar, group.by = "resolution_0.5", label = T, label.box = T, label.size = 6)+
    scale_colour_d3("category20")+scale_fill_d3("category20")
p
ggsave(plot = p, filename = "catcar_resolution_0.5.png", dpi=300, height = 6, width = 8)
```

### for my own curiosity, running dea between the two major product clusters
```{r}
Idents(catcar) <- "resolution_0.5"
product_degs <- FindMarkers(catcar, ident.1 = "0", ident.2 = "4", thresh.use = 1, min.pct = 0.25, only.pos = F, test.use = "wilcox")
product_degs <- filter(product_degs, p_val_adj < 0.05)
```

```{r}
DotPlot(catcar, group.by = "resolution_0.5", features = c("CD8A", "CD4", "CCR7", "SELL", "IL7R", "LRRN3", "TCF7", "LEF1", "GZMB", "GZMK", "GZMH", "LTB", "CD27", "CD28", "FOXP3", "TIGIT"))+scale_colour_viridis(option = "rocket", direction = -1)
```

```{r}
p <- Proportion_Plot(catcar, group.by = "resolution_0.5")+scale_fill_d3("category20")
p
ggsave(plot=p, filename = "resolution_0.5_proportions.png", dpi = 300, height = 5, width = 5)
```

```{r}
clusters <- c("0" = "c0", "1" = "c1", "2" = "c2", "3" = "c3", "4" = "c0", "5" = "c4", "6" = "c1", "7" = "c5", "8" = "c0", "9" = "c1", "10" = "c0", "11" = "c0", "12" = "c1")
Idents(catcar) <- "resolution_0.5"
catcar <- RenameIdents(catcar, clusters)
catcar@active.ident
catcar[["basic_clusters"]] <- catcar@active.ident
```

# SAVE/LOAD POINT
```{r}
#saveRDS(catcar, file = "car_only_2nd_clustering.rds", compress = F)
catcar <- readRDS("car_only_2nd_clustering.rds")
```

```{r}
p <- DimPlot(catcar, group.by = "basic_clusters", label = T, label.box = T, label.size = 6)+
    scale_colour_d3("category20")+scale_fill_d3("category20")
p
ggsave(plot = p, filename = "catcar_basic_labels.png", dpi=300, height = 6, width = 8)
```

### PSEUDOBULK DIFFERENTIAL EXPRESSION ANALYSIS
### comparing timepoints of product first
```{r}
Idents(catcar) <- "Timepoint_bin"
bulk_catcar <- AggregateExpression(catcar, assays="RNA", return.seurat=T, group.by = c("Patient", "Timepoint_bin"))
colnames(bulk_catcar)
```

### aggregating messes with metadata columns, so adding back in
```{r}
meta <- data.frame(sample = colnames(bulk_catcar))
meta <- separate(
    meta,
    col = sample,
    into = c("Patient", "Timepoint_bin"),
    sep = "_"
)
rownames(meta) <- colnames(bulk_catcar)
bulk_catcar <- AddMetaData(
    bulk_catcar,
    metadata = meta
)
```

### set up deseq dataset
```{r}
counts <- GetAssayData(
    bulk_catcar,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = bulk_catcar@meta.data,
    design = ~ Patient + Timepoint_bin
)
```


### first running LRT test just to identify genes different at any point over time
```{r}
dds <- DESeq(dds, test = "LRT", reduced = ~ Patient)
res_lrt <- results(dds)
res_lrt_df <- as.data.frame(res_lrt)
sig_lrt <- subset(
    res_lrt_df,
    padj < 0.05
)
```

### im also going to put the overall deg list into a dotplot, then extract the data to get the percentage positive values
### if a gene isn't expressed in at least 25% (subject to chnge) of cells in one timepoint, i'll exclude it as it's likely not that impactful
### ill then the updated list forward
```{r}
all_degs <- rownames(sig_lrt)
pct_exp <- DotPlot(catcar, features = all_degs, group.by = "Timepoint_bin")
pct_exp <- pct_exp$data
pct_exp <- pct_exp[, c(2,3,4)]
pct_exp <- pivot_wider(pct_exp, names_from = "id", values_from = "pct.exp")
#testing 10% to see how many that cuts
degs_filtered <- pct_exp %>%
    filter(if_any(-features.plot, ~ . >= 25))
degs_filtered <- as.data.frame(degs_filtered)
lrt_degs_filtered <- filter(sig_lrt, rownames(sig_lrt) %in% degs_filtered$features.plot)
top25_lrt_degs <- lrt_degs_filtered %>% arrange(padj) %>% slice_head(n=25)
```


### using clustered_dotplot from the scustomise package
```{r}
p <- Clustered_DotPlot(
    catcar,
    features = rownames(top25_lrt_degs),
    group.by = "Timepoint_bin",
    k = 4,
    column_label_size = 10,
    row_label_size=12,
    cluster_ident = F,
    plot_km_elbow = FALSE
)

png(
    filename = "LRT_topdegs_clustered_dotplot.png",
    width = 1200,
    height = 2000,
    res = 300
)

draw(p)

dev.off()
```

### the major differences are between the product and all other timepoints, so i will also run deseq2 without the product cells, to specifically look for differences between early, mid and late
```{r}
Idents(bulk_catcar) <- "Timepoint_bin"
no_product <- subset(bulk_catcar, idents = c("Early", "Mid", "Late"))
counts <- GetAssayData(
    no_product,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = no_product@meta.data,
    design = ~ Patient + Timepoint_bin
)
```

```{r}
dds <- DESeq(dds, test = "LRT", reduced = ~ Patient)
res_lrt <- results(dds)
res_lrt_df <- as.data.frame(res_lrt)
sig_lrt <- subset(
    res_lrt_df,
    padj < 0.05
)
top50_lrt_degs <- sig_lrt %>% arrange(padj) %>% slice_head(n=50)
```

### again i will extract genes only present in a certain percentage of cells, as a number of them are very low
```{r}
all_degs <- rownames(sig_lrt)
pct_exp <- DotPlot(catcar, features = all_degs, group.by = "Timepoint_bin")
pct_exp <- pct_exp$data
pct_exp <- pct_exp[, c(2,3,4)]
pct_exp <- pivot_wider(pct_exp, names_from = "id", values_from = "pct.exp")
#testing 10% to see how many that cuts
degs_filtered <- pct_exp %>%
    filter(if_any(-features.plot, ~ . >= 20))
degs_filtered <- as.data.frame(degs_filtered)
lrt_degs_filtered <- filter(sig_lrt, rownames(sig_lrt) %in% degs_filtered$features.plot)
top25_lrt_degs <- lrt_degs_filtered %>% arrange(padj) %>% slice_head(n=25)
```

```{r}
catcar$Timepoint_bin <- factor(catcar$Timepoint_bin, levels = c("Product", "Early", "Mid", "Late"))
Idents(catcar) <- "Timepoint_bin"
p <- Clustered_DotPlot(
    catcar,
    features = rownames(top25_lrt_degs),
    group.by = "Timepoint_bin",
    k = 5,
    column_label_size = 10,
    row_label_size=12,
    cluster_ident = F,
    plot_km_elbow = F
)

png(
    filename = "no_product_topdegs_clustered_dotplot.png",
    width = 1200,
    height = 2000,
    res = 300
)

draw(p)

dev.off()
```


### run GSEA on thye no_product LRT degs

```{r}
genes <- rownames(lrt_degs_filtered)

go_enrich <- enrichGO(
    gene = genes,
    OrgDb = org.Hs.eg.db,
    keyType = "SYMBOL"
)
```

```{r}
go_enrich
```

### no sig enriched terms, so running on original 244 genes (so sig genes but no % expressed threshold)
```{r}
genes <- rownames(sig_lrt)

go_enrich <- enrichGO(
    gene = genes,
    OrgDb = org.Hs.eg.db,
    keyType = "SYMBOL"
)
```

```{r}
enriched_pathways <- go_enrich@result
enriched_pathways <- filter(enriched_pathways, qvalue < 0.05)
```

### now running specific contrasts (for example late vs rest) to get specific signature at a timepoint
### this uses the default wald test
### still excluding product, as the initial LRT told us that this has a strong effect

```{r}
Idents(bulk_catcar) <- "Timepoint_bin"
no_product <- subset(bulk_catcar, idents = c("Early", "Mid", "Late"))
counts <- GetAssayData(
    no_product,
    slot = "counts"
)

dds <- DESeqDataSetFromMatrix(
    countData = counts,
    colData = no_product@meta.data,
    design = ~ Patient + Timepoint_bin
)

dds <- DESeq(dds)

resultsNames(dds)
```

```{r}
res_late_vs_early <- as.data.frame(results(
    dds,
    contrast = c("Timepoint_bin", "Late", "Early"))
)
sig_late_vs_early <- filter(res_late_vs_early, padj < 0.05)
```

```{r}
res_late_vs_early$group <- ifelse(res_late_vs_early$log2FoldChange > 0 & res_late_vs_early$pad < 0.05, "Increased_late", ifelse(
    res_late_vs_early$log2FoldChange < 0 & res_late_vs_early$pad < 0.05, "Increased_early", "Non_significant"))
```

```{r}
res_late_vs_early <- res_late_vs_early %>% filter(!is.na(padj))
res_late_vs_early$gene <- rownames(res_late_vs_early)
```

### get top and bottom genes for labelling
```{r}
top_late <- res_late_vs_early %>%
  slice_max(stat, n = 10)

top_early <- res_late_vs_early %>%
  slice_min(stat, n = 10)

label_genes <- bind_rows(top_late, top_early)
```

```{r}
p <- ggplot(res_late_vs_early, aes(x = log2FoldChange, y = -log10(padj), colour = group))+
    geom_point()+theme_bw()+
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", colour = "black")+
    scale_colour_manual(values = c("forestgreen", "orange", "grey50"))+
    geom_text_repel(data = label_genes,aes(label = gene),size = 3)
p
ggsave(plot = p, filename = "volcano_plot_late_vs_early.png", dpi = 300, height = 5, width = 6)
```

### GSEA
```{r}
res_late_vs_early <- res_late_vs_early[!is.na(res_late_vs_early$stat), ]
gene_list <- res_late_vs_early$stat
names(gene_list) <- rownames(res_late_vs_early)
gene_list <- sort(gene_list, decreasing = T)
```

```{r}
library(BiocParallel)
register(SerialParam())
```

### GO
```{r}
go_res <- gseGO(
    geneList = gene_list,
    OrgDb = org.Hs.eg.db,
    keyType = "SYMBOL",
    ont = "BP",
    minGSSize = 10,
    maxGSSize = 500,
    pvalueCutoff = 0.05,
    verbose = TRUE
)

go_res <- simplify(
  go_res,
  cutoff = 0.7,
  by = "p.adjust",
  select_fun = min,
  measure = "Wang",
  semData = NULL
)
```

```{r}
go_df <- as.data.frame(go_res)
```

```{r}
dotplot(go_res, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

```{r}
top_bottom_df <- go_df %>%
    arrange(desc(NES)) %>%
    slice(c(1:10, (n() - 9):n()))
```
```{r}
ggplot(top_bottom_df, aes(x= NES, y = reorder(Description, NES)))+geom_col()+theme_bw()
```

### KEGG
```{r}
# Create dataframe from ranked vector
gene_df <- data.frame(
    SYMBOL = names(gene_list),
    stat = gene_list
)

# Convert symbols to Entrez IDs
gene_conv <- bitr(
    gene_df$SYMBOL,
    fromType = "SYMBOL",
    toType = "ENTREZID",
    OrgDb = org.Hs.eg.db
)

# Merge back with statistics
gene_df <- merge(
    gene_df,
    gene_conv,
    by = "SYMBOL"
)

# Remove duplicated Entrez IDs
gene_df <- gene_df[!duplicated(gene_df$ENTREZID), ]

# Create ranked vector for KEGG
kegg_gene_list <- gene_df$stat
names(kegg_gene_list) <- gene_df$ENTREZID

# Sort decreasing
kegg_gene_list <- sort(kegg_gene_list, decreasing = TRUE)
```

```{r}
kegg_res <- gseKEGG(geneList = kegg_gene_list, organism = "hsa")
kegg_df <- as.data.frame(kegg_res)
```

```{r}
dotplot(kegg_res, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

```{r}
top_bottom_df <- kegg_df %>%
    arrange(desc(NES)) %>%
    slice(c(1:10, (n() - 9):n()))
```
```{r}
ggplot(top_bottom_df, aes(x= NES, y = reorder(Description, NES)))+geom_col()+theme_bw()
```

### REACTOME
### use kegg list (entrez id rather than synbol) as input
```{r}
reactome_res <- gsePathway(kegg_gene_list, 
                pvalueCutoff = 0.05,
                pAdjustMethod = "BH", 
                verbose = T)
reactome_df <- as.data.frame(reactome_res)
```

### get version with gene names
```{r}
reactome_res_readable <- setReadable(
    reactome_res,
    OrgDb = org.Hs.eg.db,
    keyType = "ENTREZID"
)
reactome_df <- as.data.frame(reactome_res_readable)
```

### do some plotting
```{r}
dotplot(reactome_res, showCategory=10, split=".sign") + facet_grid(.~.sign)
```

```{r}
viewPathway("Regulation of endogenous retroelements by KRAB-ZFP proteins", 
            readable = TRUE, 
            foldChange = kegg_gene_list)
```

```{r}
reactome_res$Description
```
```{r}
gseaplot2(reactome_res, geneSetID = 18, title = reactome_res$Description[18])
```

### get the genes that contribute to the KRAB-ZFP pathway
### filter on whether they are major contributors
```{r}
rank_table <- gsearank(reactome_res_readable, 18, output = "table")
major_table <- filter(rank_table, `core enrichment` == "YES")
krab_zfp_genes <- major_table$gene
```

### plot these genes in the seurat object
```{r}
DotPlot(catcar, features = krab_zfp_genes, group.by = "Timepoint_bin", scale.by = "size")+scale_colour_viridis(option = "rocket")+coord_flip()
```

```{r}
DotPlot(catcar, features = krab_zfp_genes, group.by = "time_outcome", scale.by = "size")+scale_colour_viridis(option = "rocket")+coord_flip()
```

### OVER-ENRICHMENT ANALYSIS
### get up and down degs only from the dataframe used for the volcano
```{r}
up_late <- filter(res_late_vs_early, group %in% "Increased_late")
up_late <- rownames(up_late)


go_enrich <- enrichGO(gene = up_late,
                      universe = names(rownames(res_late_vs_early)),
                      OrgDb = org.Hs.eg.db,
                      keyType = 'SYMBOL',
                      ont = "BP",
                      pvalueCutoff = 0.05, 
                      qvalueCutoff = 0.10)
go_enrich <- as.data.frame(go_enrich)
```